---
title: "SumbarajuAditya_ex1_2_R_Refresher"
author: "SumbarajuAditya"
date: "8/30/2021"
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown



```{r}
# Load Libraries

library(blsAPI)
library(dplyr)
library(ggplot2)
```

# Load BLS data

```{r}
#Consumer Price Index-All Urban Consumers (Current Series) 

urban_cpi <- blsAPI('CUUR0000SA0', return_data_frame = TRUE)
# Chained CPI-All Urban Consumers
chained_cpi <- blsAPI('SUUR0000SA0', return_data_frame = TRUE)


```

# check for right grain of data

```{r}
summary(urban_cpi)
summary(chained_cpi)

str(urban_cpi)
str(chained_cpi)

dim(urban_cpi)
dim(chained_cpi)

```
# Data cleansing and Merging Step 

```{r}
#Merging two datasets
df1 <- merge(urban_cpi, chained_cpi, by=c("year","period"))

#Renaming the columns
names(df1)[names(df1) == "value.x"] <- "urban_cpi_Rate"
names(df1)[names(df1) == "value.y"] <- "chained_cpi_Rate"

#Convert string representations to numbers
df1$urban_cpi_Rate <- as.numeric(df1$urban_cpi_Rate)
df1$chained_cpi_Rate <- as.numeric(df1$chained_cpi_Rate)

summary(df1)
```
#Histograms of the 2 variables

```{r}
ggplot(df1, aes(x = urban_cpi_Rate)) + geom_histogram (col = c("blue"))
ggplot(df1, aes(x = chained_cpi_Rate)) + geom_histogram(col = c("green"))

```
#Boxplots of the 2 variables

```{r}
ggplot(df1, aes(x = urban_cpi_Rate)) + geom_boxplot(col = c("blue"))
ggplot(df1, aes(x = chained_cpi_Rate)) + geom_boxplot(col = c("green"))
```

#Density plots of the 2 variables

```{r}
ggplot(df1, aes(x = urban_cpi_Rate)) + geom_density(col = c("blue"))
ggplot(df1, aes(x = chained_cpi_Rate)) + geom_density(col = c("green"))

```

#Save the data-frame as a CSV file

```{r}
write.csv(df1, 'C:/BU/DSC630/Consumer_Price_Index.csv')
```


### 2. Bivariate plots and correlation

```{r correlation, echo=FALSE}
#Bivariate plots
ggplot(df1, aes(x = urban_cpi_Rate, y = chained_cpi_Rate)) + geom_point(col = c("red"))
# Let's calculate the correlation using pearson method
library(tidyverse)
cor(df1$urban_cpi_Rate, df1$chained_cpi_Rate, method = "pearson")

#Cross tables
library(gmodels)
CrossTable(df1$urban_cpi_Rate, df1$chained_cpi_Rate)
```


### 3. Organize a Data Report
```{r}
summary(df1)

str(df1)

dim(df1)

```

## Discussion points
1. The median of urban_cpi_Rate is 258 where as chained_cpi_Rate stands 144.8
2. Bivariate plots resembles the positive correlation.
3. The pearson correlation coefficient is calculated as 0.9981076. it means every positive increase in urban_cpi_Rate there is a positive increase in chained_cpi_Rate
4. The box plots depicts the positive skewness.

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
